% ============================================================================
% Abstract for EVGA2021
% ============================================================================

\documentclass[natbib,onecolumn,twoside]{svmultiag}

\usepackage{url}

% ============================================================================

\title*{First results from the new station NYALE13S}

\author{A-S.~Kirkvik}

\institute{
Ann-Silje Kirkvik$^1$
\at
(1) Norwegian Mapping Authority
Kartverksveien 21, 3511 H{\o}nefoss, Norway\\
Corresponding author: \url{ann-silje.kirkvik@kartverket.no}          
}
% ============================================================================
\begin{document}
\maketitle
\thispagestyle{empty}
% ----------------------------------------------------------------------------
\abstract{
In February 2020 the new VLBI station NYALE13S made its first successful 24 hour session.
The station is equipped with a tri-band receiver and will participate in R1 and R4 sessions 
together with NYALES20 before NYALES20 is dismantled and NYALE13S is equipped with a 
broadband receiver. The goal is to have at least one and a half year of parallel observations.
The purpose of the legacy observations is to be able to transfer the long time series from 
NYALES20 to the new station NYALE13S and compare the results with the local tie measurements.
The length of the NYALES20-NYALE13S baseline is approximately 1.5 kilometers and these parallel
observations will be very useful for connecting future VGOS observations with the legacy
network. Some early analysis results from the observations gathered so far will be presented.
The number of available observations is significantly less than scheduled due to problems with
the equipment both at the new and old station. The results are therefore very preliminary and
hopefully more sessions will be observed in the near future. 
}
% ----------------------------------------------------------------------------
\section*{Session-3: Analysis}
\section*{Presentation type: Poster}

% ----------------------------------------------------------------------------
\end{document}
% ============================================================================

